Learning From Massive Noisy Labeled Data for Image Classification

### 2015-04-06 paper.tex
Polish the section of deep learning

We estimate $\mat{C}$ on a relatively small dataset $\mathcal{D}_c = \{(\vec{x},\vec{y},\tilde{\vec{y}})\}_N$, where we have $N$ images with both clean and noisy labels. As prior information about $\vec{z}$ is not available, we just solve the following optimization problem:
\begin{equation}
    \max_{\mat{C},\vec{z}^{(1)},\cdots,\vec{z}^{(N)}} \sum_{i=1}^N \log p(\tilde{\vec{y}}^{(i)}|\vec{y}^{(i)},\vec{z}^{(i)})
\end{equation}
Obviously, sample $i$ contributes nothing to the optimal $\mat{C}^*$ if $\vec{y}^{(i)}$ and $\tilde{\vec{y}}^{(i)}$ are equal. So that we ignore those samples and reinterpret the problem in another form by exploiting Eq~\ref{eq:likelihood}:
\begin{equation}
\begin{aligned}
\max_{\mat{C},\vec{t}} \quad & E=\sum_{i=1}^{N'} \log \alpha^{\vec{t}_i} + \log (\tilde{\vec{y}}^{(i)T} \mat{C} \vec{y}^{(i)})^{1-\vec{t}_i}\\
\text{subject to} \quad & \mat{C} \text{ is a stochastic matrix of size } L\times L \\
                    & \vec{t} \in \{0,1\}^{N'}
\end{aligned}
\end{equation}
where $\alpha=\frac{1}{L-1}$ and $N'$ is the number of remaining samples. The semantic meaning of the above formulation is that we need to assign each $(\vec{y},\tilde{\vec{y}})$ pair the optimal noise type, while finding the optimal $\mat{C}$ simultaneously.

@@

We integrate the proposed label noise model into a deep learning framework. As demonstrated in Figure~\ref{fig:framework}, we predict the probability $p(\vec{y}|\vec{x})$ and $p(\vec{z}|\vec{x})$ by using two independent CNNs. Moreover, we append a label-noise-model layer at the end, which takes as input the CNNs' prediction scores and the observed noisy label. Stochastic Gradient Ascent (SGA) with backpropagation technique is used to approximately optimize the whole network. In each forward pass, the label-noise-model layer computes the posterior of latent variables according to Eq~\eqref{eq:posterior}. While in the backward pass, it computes the gradients according to Eq~\eqref{eq:gradient}.

@@

Directly training the whole network with random initialization is impractical, because the posterior computation could be totally wrong. Therefore, we need to pretrain each CNN component with strongly supervised data. Gathering ground truth object classes is straight forward, since we can just manually label some images by expert. The resulting strongly labeled dataset $\mathcal{D}_c$ can be directly used for training the network that predicts $p(\vec{y}|\vec{x})$. On the other hand, although off-the-shelf supervision for $p(\vec{z}|\vec{x})$ is not available, we can heuristically generate some data by utilizing the images having both strong and noisy labels. For each sample, we choose as ground truth the $\vec{z}$ that maximizes the likelihood in Eq~\eqref{eq:likelihood}.

After both the CNN components are properly pretrained, we can start training the whole network with massive noisy labeled data. However, some practical issues need to be further discussed. First, if we merely use noisy labels, we will lose precious knowledge that we have gained before and the model could be drifted. Therefore, we need to mix strongly label data together in to our training set, which is depicted in Figure~\ref{fig:framework} as the extra supervisions for the two CNN components. Then each CNN receives two kinds of gradients, one is from the clean labeled data and the other is from the noisy labeled data. We denote them by $\Delta_c$ and $\Delta_n$, respectively. A potential problem is that $|\Delta_c| \ll |\Delta_n|$, because clean data is much less than the noisy data. To deal with this problem, we bootstrap the clean data $\mathcal{D}_c$ to half amount of the noisy data $\mathcal{D}_n$. This upsampling process brings another advantage --- the gradients we calculated in each mini-batch are much more stable.

Our proposed method has the ability to figure out the ground truth label given the image and its noisy label. From the perspective of information, our model predicts from two kinds of clues: what are the true labels for other similar images; and how confusing is the input image itself. Label Propagation method~\cite{zhu2002learning} explicitly uses the first kind of information, while we implicitly capture it with a discriminative deep model. Meanwhile, we exploit the second kind of information to bridge the semantic gap between the image and its possible noisy labels.
% section weak_label_deep_learning (end)

>>>>

We estimate $\mat{C}$ on a relatively small dataset $\mathcal{D}_c = \{(\vec{x},\vec{y},\tilde{\vec{y}})\}_N$, where we have $N$ images with both clean and noisy labels. As prior information about $\vec{z}$ is not available, we just solve the following optimization problem:
\begin{equation} \label{eq:opt_all}
    \max_{\mat{C},\vec{z}^{(1)},\cdots,\vec{z}^{(N)}} \sum_{i=1}^N \log p(\tilde{\vec{y}}^{(i)}|\vec{y}^{(i)},\vec{z}^{(i)})
\end{equation}
Obviously, sample $i$ contributes nothing to the optimal $\mat{C}^*$ if $\vec{y}^{(i)}$ and $\tilde{\vec{y}}^{(i)}$ are equal. So that we ignore those samples and reinterpret the problem in another form by exploiting Eq~\ref{eq:likelihood}:
\begin{equation}
\begin{aligned}
\max_{\mat{C},\vec{t}} \quad & E=\sum_{i=1}^{N'} \log \alpha^{\vec{t}_i} + \log (\tilde{\vec{y}}^{(i)T} \mat{C} \vec{y}^{(i)})^{1-\vec{t}_i}\\
\text{subject to} \quad & \mat{C} \text{ is a stochastic matrix of size } L\times L \\
                    & \vec{t} \in \{0,1\}^{N'}
\end{aligned}
\end{equation}
where $\alpha=\frac{1}{L-1}$ and $N'$ is the number of remaining samples. The semantic meaning of the above formulation is that we need to assign each $(\vec{y},\tilde{\vec{y}})$ pair the optimal noise type, while finding the optimal $\mat{C}$ simultaneously.

@@

We integrate the proposed label noise model into a deep learning framework. As demonstrated in Figure~\ref{fig:framework}, we predict the probability $p(\vec{y}|\vec{x})$ and $p(\vec{z}|\vec{x})$ by using two independent CNNs. Moreover, we append a label noise model layer at the end, which takes as input the CNNs' prediction scores and the observed noisy label. Stochastic Gradient Ascent with backpropagation technique is used to approximately optimize the whole network. In each forward pass, the label noise model layer computes the posterior of latent variables according to Eq~\eqref{eq:posterior}. While in the backward pass, it computes the gradients according to Eq~\eqref{eq:gradient}.

@@

Directly training the whole network with random initialization is impractical, because the posterior computation could be totally wrong. Therefore, we need to pretrain each CNN component with strongly supervised data. Images and their ground truth labels in the dataset $\mathcal{D}_c$ are used to train the CNN that predicts $p(\vec{y}|\vec{x})$. On the other hand, the optimal solutions of $\vec{z}^{(1)},\cdots,\vec{z}^{(N)}$ in Eq~\eqref{eq:opt_all} are used to train the CNN that predicts $p(\vec{z}|\vec{x})$.

After both CNN components are properly pretrained, we can start training the whole network with massive noisy labeled data. However, some practical issues need to be further discussed. First, if we merely use noisy labels, we will lose precious knowledge that we have gained before and the model could be drifted. Therefore, we need to mix strongly label data into our training set, which is depicted in Figure~\ref{fig:framework} as the extra supervisions for the two CNNs. Then each CNN receives two kinds of gradients, one is from the clean labeled data and the other is from the noisy labeled data. We denote them by $\Delta_c$ and $\Delta_n$, respectively. A potential problem is that $|\Delta_c| \ll |\Delta_n|$, because clean data is much less than the noisy data. To deal with this problem, we bootstrap the clean data $\mathcal{D}_c$ to half amount of the noisy data $\mathcal{D}_n$. This upsampling process brings another advantage --- the gradients we calculated in each mini-batch are much more stable.

Our proposed method has the ability to figure out the true label given the image and its noisy label. From the perspective of information, our model predicts from two kinds of clues: what are the true labels for other similar images; and how easily for the image to be mislabeled. Label Propagation method~\cite{zhu2002learning} explicitly uses the first kind of information, while we implicitly capture it with a discriminative deep model. On the other hand, the second kind of information correlates the image content with the label noise, which can help distinguish between hard samples and mislabeled ones.
% section weak_label_deep_learning (end)
